h2. Stats

Each building has some factors which represent its impact on the environment, its efficiency and attractivity.

h5. Population

!>img/pop.png! Shows how many people can live in a building or, if the number is negative, it shows how many people can work in that building. If you don't have enough workers, your income and production is reduced.

h5. Energy

!>img/energy.png! Energy represents how much energy does a building consume or produce. If you don't produce enough energy, you can't build more buildings.

h5. Pollution

!>img/pol.png! Pollution shows how ecological your town is. It should be as low as possible.

h5. Garbage

!>img/waste.png! Buildings produce trash. This numbers shows how much tons of trash does a building produce or, if it is negative, it represents how many tons can it recicle. For the efficiency of your town, you should try to recicle the majority of the garbage you produce.

h5. Income

!>img/income.png! The income represents how much money you gain from the building (by taxes) or how much does it cost you to keep the building up and running. Goods are needed in order to build new buildings.

h5. Production

!>img/prod.png! The production represents how much goods does your building produce or consume. You need goods in order to build complex buildings.

h5. Water

!>img/water.png! Represents how much water does a building consume. Water is provided by water storage, production or treatment facilities and it's used in the other buildings.

h5. Score

!>img/score.png! Shows how much points you gain by building this particular building. You might not gain exacly this ammount because your score is also affecter by other factors which multiply or reduce it.